---
title: "Introduction to darthpack"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{aa-introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography:
- book.bib
- packages.bib  
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE, warning=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'dplyr', 'truncnorm', 
                           'lhs', 'IMIS', 'matrixStats',
                           'plotrix', 'psych', "survival",
                           "scatterplot3d", "reshape2",
                           "BiocManager", "devtools",
                           "shiny", "rstudioapi", 'nlme'
), 'packages.bib')
```

## Set-up {-}
The vignettes in the *vignettes* folder and the *Articles* tab in the [`darthpack` website](https://darth-git.github.io/darthpack/) are meant to guide you through the `R` code-based functions of `darthpack` with a fully functional decision model to showcase the framework described by the [Decision Analysis in R for Technologies in Health (DARTH) workgroup](http://darthworkgroup.com/) in the manuscript *A need for change! A coding framework for improving transparency in decision modeling* [@Alarid-Escudero2019b]. For a more detailed description of the usage and installation of `darthpack`, we refer the reader to the package's website (https://darth-git.github.io/darthpack/). The framework is divided into different directories, described in Table 1 of the main manuscript, that could be accessed from the RStudio project *darthpack.Rproj*. We refer to the directory names of this framework and scripts stored in these directories using *italic* style. These vignettes are created with R Markdown and are located in the *vignettes* directory of the framework. The figures for the case-study can be found in the *figs* directory, data required to conduct some of the analyses of the different components are in the *data* and *data-raw* directories and the `R` scripts with functions, are located in the *R* directory. The main `R` scripts that conduct the analyses of the different components of the framework are stored in the `analysis` directory. In this document we do not show all the `R` code we refer to. Therefore, it is important to follow along while reading this document. To make sure you have all the required packages needed to run all the code, first install all the required and suggested packages in *DESCRIPTION* file in the main directory of `dartpack`.

## Shiny app

## Prerequisites general{-}

The book describing our Sick-Sicker case study is written in R Markdown using the `bookdown` package. More about books written in R Markdown using `bookdown` can be found on this website: <https://bookdown.org/yihui/bookdown/>. When you like to generate the book yourself using Markdown you need to install the `bookdown` package from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

# References {-}